---
title: "R Notebook"
output: html_notebook
---
Loading libraries

```{r}
library(tidyverse)
```
load the dataset


```{r}
data <- read.csv("heart.csv")
str(data)
```
```{r}
nrow(data)
```


```{r}
ncol(data)
```


```{r}
colnames(data)
```
Data Transformation
```{r}
data2 <- data %>%
  mutate(sex = if_else(sex == 1, "MALE", "FEMALE"),
         fbs = if_else(fbs == 1, ">120", "<=120"),
         exang = if_else(exang == 1, "YES" ,"NO"),
         cp = if_else(cp == 1, "ATYPICAL ANGINA",
                      if_else(cp == 2, "NON-ANGINAL PAIN", "ASYMPTOMATIC")),
         restecg = if_else(restecg == 0, "NORMAL",
                           if_else(restecg == 1, "ABNORMALITY", "PROBABLE OR DEFINITE")),
         slope = as.factor(slope),
         ca = as.factor(ca),
         thal = as.factor(thal),
         target = if_else(target == 1, "YES", "NO")
         ) %>% 
  mutate_if(is.character, as.factor) %>% 
  dplyr::select(target, sex, fbs, exang, cp, restecg, slope, ca, thal, everything())
```
```{r}
view(data2)
```


Data Visualization
```{r}
#bar plot for target (heart disease)
ggplot(data2, aes(x =data2$target, fill =data2$target))+
  geom_bar()+
  xlab("heart disease")+
  ylab("count")+
  ggtitle("Presence and absence of heart disease")+
  scale_fill_discrete(name='Heart Disease', labels = c("Absense", "Presence"))
```




```{r}
#proportion of the people who have and do not have heart disease in decimals
prop.table(table(data2$target))
```


```{r}
# count the frequency of the values of age

data2 %>%
  group_by(age) %>%
  count() %>%
  filter(n>10) %>%
  ggplot()+
  geom_col(aes(age, n), fill = 'blue')+
  ggtitle("Age Analysis")+
  xlab("Age")+
  ylab("Frequency")
```
compare blood pressure across chest pain
```{r}
data2 %>%
  ggplot(aes(x=sex, y=trestbps))+
  geom_boxplot(fill ='cyan')+
  xlab('sex')+
  ylab('BP')+
  facet_grid(~cp)
```
see what we get when we plot the un mutated data
```{r}
data %>%
  ggplot(aes(x=sex, y=trestbps))+
  geom_boxplot(fill ='cyan')+
  xlab('sex')+
  ylab('BP')+
  facet_grid(~cp)
```
Box plots with respect to cholestorel


```{r}
data2 %>%
  ggplot(aes(x=sex, y=chol))+
  geom_boxplot(fill ='orange')+
  xlab('sex')+
  ylab('Chol')+
  facet_grid(~cp)
```
Lets work with some correlation now
```{r}
install.packages("corrplot")
```


```{r}
library(corrplot)
library(ggplot2)
```
```{r}
cor_heart <- cor(data2[, 10:14])
cor_heart

corrplot(cor_heart, method ='square', type='upper')
```


the larger reddish squares indicate larger negative correlation and larger blue squares indicate strong positive correlation
e.g: age and tesbps have a strong positive correlation while age and thalach have strong negative correlation

